How to Analyze Build Logs --- A~Comparative Study of Chunk~Retrieval~Techniques

### 2020-04-22 tse-paper/content.tex
First feedback and review integrated

Within the past two decades, Continuous Integration (CI) has become a
ubiquitous best practice to streamline the build process of software
projects~\cite{hilton2016usage,beller2017oops,staahl2014modeling}.
Build
logs are a textual by-product that automatic software build processes
create.
As a ``treasure trove'' of information, they contain results
not only of the elementary steps to ``make'' a project---such as
retrieving and resolving dependencies, compiling, and testing the
software---but also about ancillary quality assurance steps such as
automated static analysis tools.

When a build or one of its steps fails, developers typically
scavenge the log for the source of the error.
This manual activity is
tedious and prone to error~\cite{santolucito2018statically}.
As a
single build log can be over 50 megabytes large~\cite{beller2017oops},
finding a piece of information in it is akin to searching for a
needle in a haystack.
In addition, due to their information richness,
build logs enable a plethora of other applications: with the help of
build logs, we can better understand and categorize CI
failures and compile
errors~\cite{islam2017insights,seo2014programmers}, differentiate
testing practices~\cite{orellana2017differences,vassallo2017a-tale},
train classifiers to predict the build outcome and
duration~\cite{ni2017cost,bisong2017built,machalica2019predictive}, or
investigate static analysis tools in CI
builds~\cite{zampetti2017open}.
Manual analysis of build logs does not
scale to the number of build logs and amount of information that such
sophisticated applications require.
Therefore, being able to
automatically extract pieces of information from build logs
is not only paramount to applications such as the above mentioned
ones, but can also help
developers debug a broken build.

@@

Despite its paramount role, the process of analyzing
build logs has so far received little attention, bringing us to our
first research question:
\begin{simplebox}[minipage boxed title*=-5cm]{\textbf{Research Question
1}}
What are the state-of-the-research techniques for build log analysis?
\end{simplebox}

In this article, as the first step (\Cref{fig:overview}), we show that
less than half of the 61 works which rely on build log analysis
describe how they performed it.
Even fewer did so in detail or
made their implementation available.
Through the literature
survey we identify three prevalent strategies to
analyze build logs: Using 1) a set of
manually curated regular expressions,
similar to TravisTorrent~\cite{beller2017oops},
2) information
retrieval techniques,
and 3) a list of search keywords
to identify regions of interest in the build log.

Even though these three approaches to retrieve information chunks from
build logs have different strengths and weaknesses, the literature
survey shows that their use is often arbitrary and that there is
usually no discussion on why to use one over the other.
Developers and
researchers not only seem to lack awareness of which techniques exist
(RQ1), but also guidelines on when to employ which.
As a consequence,
in RQ2 we ask:

@@

To answer this question, we implement three automatic techniques for
build log analysis based on the identified strategies (step 2):

\begin{itemize}
  \item \textbf{Program Synthesis by Example (referred to as PBE)}
  Based on user examples of build logs and chunks, PBE
  synthesizes
  a regular expression matching the given chunks within the build logs.
  \item \textbf{Common Text Similarity (CTS)}
  Using the Vector Space Model, CTS selects those lines of a log which are
  most similar to the lines present in example chunks provided by
  the user.
  \item \textbf{Keyword Search (KWS)}
  Ad-hoc method for finding fitting passages by searching the whole
  text for
  the occurrence of specific trigger words.
\end{itemize}

In step 3, we conduct an empirical study to gauge the performance of
these chunk retrieval techniques in
real-world projects under a variety of conditions.
Our study is based on the \emph{LogChunks} data
set~\cite{brandt2020logchunks} which encompasses build logs from
80 popular open-source projects and a broad range of
build environments.
Finally, we develop guidelines (step 4) on when
one of the techniques might be preferential to its alternatives.

We recommend PBE for use cases where the desired information is always
represented in the same structural way and high confidence in
precision and recall of the chunk retrieval is required.
Results
produced by PBE are suited for automatic on-ward processing.
CTS is
well suited when the representation of the desired information varies
slightly and the output of the chunk retrieval is further processed by
a human.
In cases where the textual representation of the desired
information in the log is unpredictable or varies greatly, KWS seems
to be the best choice.
However, its low precision---it extracts a
context of multiple lines around a finding---makes it generally
unsuited for automatic on-ward processing and instead requires a human
to further inspect and interpret the output chunk.

To enable the retracing of our research, we provide a detailed replication
package~\cite{brandt2020chunk-replication}.

@@

The field of analyzing build
logs is relatively young; We have seen scattered
research efforts on build log analysis.
However, the processing of build logs was often
a side-part of the article.
To the best of our
knowledge, there has not been an attempt to systematically overview
and survey approaches for processing build logs.
We conducted a systematic mapping study to determine the
state-of-research on build log analysis following the guidelines
by Petersen et al.~\cite{petersen2008systematic,petersen2015guidelines}.

In the following section, we explain how build log analysis differs
from system log analysis.
Further, we describe how we selected the articles for our literature
survey and the data extraction from the resulting articles.
We present our results on how build logs are analyzed in literature
and discuss the implications of these findings on our research.

@@

\subsection{Distinction from System Log Analysis}
\label{sec:system-log-analysis}
A related field of build log analysis is the processing of system log
files produced during runtime.
The main difference between build logs
and system logs is that system logs are fundamentally structured
through events.
Each line in a log file represents one event with a
set of fields: timestamp, verbosity level and raw message
content~\cite{he2017towards}.
\Cref{lst:system-log} shows
example lines from a system log.

The first goal in analyzing system log files is ubiquitously to
separate constant and variable parts within a log
message~\cite{nagappan2010abstracting,he2017towards}.
Next, the log
messages are clustered into log events, unifying messages with
identical constant parts and varying parameters.
The output of a log
parser is an ordered list of timed events and their corresponding
parameter values~\cite{he2016evaluation}.
This structured log is then
the input to various machine learning and data mining processes.

The techniques developed for system log analysis can also be applied
to build logs.
One example is comparing execution traces to reference
traces of intended behavior to detect anomalies.
Amar et
al.~\cite{amar2019mining} employed a similar approach to detect
relevant lines in build logs.
However, many of the techniques developed to analyze system logs
leverage their inherent event structure.
As build logs lack this structure, we focussed our literature
survey on articles that as a source of information.
% In this article, we focus on extracting a single specified information
% from the build log as a whole with chunk retrieval techniques.
% Chunk
% retrieval techniques are used as a part of log parsing to retrieve the
% values of variable parts in a log message, e.g.\ by using regular
% expressions~\cite{nagappan2010abstracting,xu2009detecting}.

\subsection{Study Selection}
\Cref{fig:lit-survey} presents an overview of our study selection
process.
Since the field of build log analysis is so young, we
surveyed a population of scientific material as broad as possible.
We did not limit our sources to specific Software Engineering venues, as
is often done~\cite{petersen2015guidelines}, but survey all sources
monitored by Google Scholar, including Bachelor's and Master's
theses and academic slide decks.
We use Google Scholar as opposed to Scopus or
other databases, as it is the search engine for academic articles with
the broadest and most current index, including preprints.

Our Google Scholar search for variations of ``build log''
returns more than 2.2 million results (2020-03-15).
We refined our search criteria to exclude obvious
non-related works.
For example, we decided to exclude systems
or event logs (see \Cref{sec:system-log-analysis}).
This left us with 1,552 search results
(2020-03-15).
The full search queries are documented in our replication
package~\cite{brandt2020chunk-replication}.

To make handling so many search results feasible, we employed a
two-pass filtering strategy:

First, we filtered articles based on (1) their title and (2) the
contextual information Google Scholar displayed.
We included only
publications written in a language intelligible to the authors (that
is, English or German), which disregarded fewer than 1\% of search
hits.
We then unified
the results of the four search queries based on the link as the
identifying element, removing 60 articles that appeared in more than
one search query.
We removed further duplicates based on the title and were
left with 256 (16\%) of the original search results.
% removed duplicates of data extraction pass already here

With these 256 remaining articles, we followed a backward snowball
sampling for related work.
If one of the articles referenced a new article in the context
of build log analysis we added the new to our literature set.
This added five papers (eight before
duplicate elimination), coming to a total of 261
articles.

Subsequently, we performed a second, finer filter phase by downloading the
full articles, reading their abstracts and searching the full text
for the occurrence of ``log''.
If these papers showed traces of
working with build logs, we included them in our survey.
In total, we
were left with 87 articles after the two filtering phases.

\subsection{Article Classification}
Having trimmed down the set of articles, we investigated them in depth.

We wanted to characterize how researchers use build logs as input to
any processes.
We were interested in
\begin{itemize}
  \item what information they are retrieving from the logs and
  what they use the retrieved information for (\textbf{RQ1.1})
  \item which technique they are using to process the build logs,
  in how much detail it is described and whether they published their
  tools (\textbf{RQ1.2})
  \item what kind of build logs they are using and the origin of these
  logs.
(\textbf{RQ1.3})
\end{itemize}
To objectify our data extraction we created a template of the targeted
factors and performed a pilot, as recommended by Petersen et
al.~\cite{petersen2015guidelines}.
Two authors extracted the data from five articles and we held a
consensus meeting to unify our understanding
of the data extraction template.
The remaining articles were divided over these two authors.
We inspected the articles as closely as necessary,
starting from occurrences of ``log'' or ``build'' within the text.

Relevant for our survey were all articles that
use logs from software builds as input to
some kind of process.
This excluded 15 articles from our study.
In several cases, multiple articles were extensions of others.
Of these, we only considered the earliest article which reported on
the processing of build logs.
This excluded another 11 articles.
In sum, we further extracted data from 61 articles.

The full data extraction template and our results of this step can be
found in our replication package~\cite{brandt2020chunk-replication}.
Note that for most questions we allowed assigning multiple values,
e.g.\ if
two kinds of information were targeted during a build log analysis.

\begin{simplebox}[attach boxed title to top center={yshift=-6mm}]
{\textbf{RQ1:} What are the state-of-the-research
techniques for build log analysis?}
\begin{itemize}[leftmargin=2cm]
  \item[\textbf{RQ1.1:}] Which information is targeted?
  \item[\textbf{RQ1.2:}] Which techniques are used?
  \item[\textbf{RQ1.3:}] Which build logs are analyzed?
\end{itemize}
\end{simplebox}

@@

\subsection{Results}
In the following, we will present the results of our literature survey.
We present the information within the build logs that was targeted,
the retrieval techniques used and the origin of the build logs
which were analyzed.

\subsubsection{Retrieved Information}
The information that was targeted varied from article to article.
Most prominent was the search for errors (39\%) which occurred within
the build process, as presented in \Cref{fig:litsur:info_target}.
This is followed by extracting the executed or the failing
tests (both 11\%).
There were also numerous cases which targeted more specialized
information, such as the environments used during the
build~\cite{zolfagharinia2017not}, packages
installed~\cite{selberg2012use} or hints to source files which are
related to build failures~\cite{ren2018automated}.

In most cases (51\%), a chunk was retrieved from the log.
Build logs were scanned for compilation
errors~\cite{clemencic2014new} or
the duration of tasks within the build~\cite{zhang2016android}.
36\% of the articles classified the build logs into categories.
For instance, distinguishing failures caused by developers from
failures caused
by infrastructure errors~\cite{lindqvist2019detection} or
whether projects use automated static analysis
tools~\cite{kavaler2019tool}.
Other types of information (e.g.\ counting
lines, summarization) were only retrieved
in one case each.

The majority of articles (46\%) used the outcome of their build log
analysis for research purposes, as shown in \Cref{fig:litsur:use}.
Further uses were giving feedback to developers (10\%) or automatic
on-ward processing of the retrieved information (8\%).

\subsubsection{Retrieval Technique}
\Cref{tab:litsur:techniques} presents the techniques that the articles
described for processing build logs.
43 of the 61 articles we inspected mentioned how they are analyzing build
logs.
However, only 16 of these described their method in detail.

The most mentioned method (34\%) was using a parser, where we also
included
rather imprecise statements like ``we parse the build logs''.
25\% of the articles described the use of regular expressions and 20\%
inspected the logs manually.
Another 8\% employed machine learning techniques such as natural language
processing or information retrieval techniques.
To give examples,
Seo et al.~\cite{seo2014programmers} developed a custom
parser to classify error messages, while Vassallo et
al.~\cite{vassallo2017a-tale} analyzed build logs with regular
expressions.
Ghaleb et al.~\cite{ghaleb2019studying} used a compound approach.
They start with manual categorization of the build logs and select
search strings that identify their targeted categories.
Based on the search strings they create a script that automatically
classifies the remaining logs.

25\% of the articles claimed their implementation is available,
however, we
saw a very low reuse of build log analysis tools within our survey.
Five articles employed the tools used to create
\emph{TravisTorrent}~\cite{beller2017travistorrent,beller2017oops,
orellana2017differences,zhao2018comparing} or
an extended version of them~\cite{rott2019empirische,
shi2018evaluating},
two used the
\emph{Maven Log Analyzer}~\cite{macho2018automatically,gallaba2018noise},
another two the
\emph{MAKAO}~\cite{wen2018blimp,adams2007design,adams2007makao} tool.

\subsubsection{Build Logs}
The distribution of supported build tools is shown in
\Cref{fig:litsur:log_producer}.
Of the articles we surveyed, 30\% analyzed build logs produced by
Maven~\cite{maven2019website}, 16\% analyzed logs produced by
Gradle~\cite{gradle2020website}
and 10\% analyzed logs produced by Ant~\cite{ant2020website}.

Sometimes this limitation to a specific build tool is motivated by
aspects separate from the build log analysis.
For example, Shi et al.~\cite{shi2018evaluating} focussed on Maven
build logs because they also chose the PIT tool to calculate coverage
and mutation score.
A larger amount of the articles within our survey, however, limit
themselves to few build tools because their log format varies and
therefore ``parsers must be specialized to each build and test
framework''~\cite{tomassi2019bugswarm}.

Only 7\% of the proposed methods claimed to be language-agnostic, while
several described they are covering multiple source build tools.
The majority of articles (42\%) targeted logs from a source that was not
targeted by any second article.

The main platform that build logs were collected from is Travis
CI~\cite{travisci2019webpage},
which was used in 38\% cases.
18\% of the articles processed logs from TravisTorrent and 8\% included
build logs from industrial projects.

The number of logs analyzed varied greatly and we were not always able to
extract it confidently.
Articles reported processing from two up to 122 million logs.

31\% of the surveyed articles claimed to have published their data to
enable further research.
However, we found no clear reuse of build log data sets with the
exception of TravisTorrent~\cite{beller2017travistorrent}, which was
analyzed by Vassallo et al.~\cite{vassallo2017a-tale},
Ghaleb et al.~\cite{ghaleb2019studying} and several other studies.

\begin{figure}[tbhp]
		\centering
		\includegraphics[width=\columnwidth, trim={1.1cm 0.4cm
		1.5cm 0.5cm},
		clip]{img/lit-sur/log_producer_annotated.pdf}
		\caption{Frequency of supported log producers.}
		\label{fig:litsur:log_producer}
\end{figure}

\subsection{Discussion}
\label{sec:lit-sur:discussion}
Our literature survey sheds light on how build log analysis is used within
computer science research.
Hereafter we will summarize our findings and relate them to the
approach we are taking to improve build log analysis with chunk
retrieval techniques.

We saw that various researchers process build logs to support their
research, often because the logs provide more or the necessary information
to the researchers~\cite{ren2018automated}.
Articles which explicitly mention a manual analysis of
the logs, mostly restrict their work to few logs or take a
``representative sample to make manual analysis
feasible''~\cite{zolfagharinia2017not}.
Automated approaches are necessary as many studies target a large
number of builds.
% The chunk retrieval techniques we investigate within this article are
% automated, apart from their configuration through training examples.

We have observed very little reuse of build log analysis tools.
This can stem from the fact that publishing replication packages and
research prototypes became popular only in recent years or from the
high specialization of the developed tools in regards to supported
build environments and targeted information.

Several articles noted the variety in log formats for different build
environments, which requires customizing build log techniques for
every supported build environment.
\Cref{fig:litsur:log_producer} shows that few build environments are
targeted by many studies and tools (A).
The majority of build environments was however only targeted by one
article each (B).
From this, we follow that, while custom parsers are beneficial for the
small group of environments targeted by many (A), all and especially
the rarer investigated environments (B) would benefit from a generic
solution.
% We implement build environment agnostic build log analysis
% techniques, which are customized to a specific project by
% the training examples which the user provides.
% In addition, the \emph{LogChunks}~\cite{brandt2020logchunks}
% data set we chose for our evaluation covers a broad range of build
% environments.
% % Further, the training examples also specify the log chunk targeted
% by the
% % retrieval, which enable our chunk retrieval techniques to retrieve
% % a broad range of information from build logs.

Urli et al.~\cite{urli2018design} strongly discouraged from parsing
build logs for information as it is ``too error prone'', other
articles pointed to the effort of developing a custom tool.
Regular expressions are known to be tedious to
maintain~\cite{michael2019regexes}.

>>>>

Within the past two decades, Continuous Integration (CI) has become a
ubiquitous best practice to streamline the build process of software
projects~\cite{hilton2016usage,beller2017oops,staahl2014modeling}.
Build logs are a textual by-product that automatic software build
processes create.  As a treasure trove of information~\cite{meyer},
build logs contain trace outputs of not only the elementary steps to
``make'' a project---such as retrieving and resolving dependencies and
compiling---but also about ancillary quality assurance steps---such as
testing the software and running automated static analysis tools. The
contents of build logs can be highly variable depending on which tools
are involved in the build process and how they are
configured. The format of logs changes as projects and build tools evolve over time. Typically, though, they are at least a semi-structured and
somewhat stable journal of the commands executed during the build and
their results.

When a build or one of its steps fails, developers typically
scavenge the logs for information related to the source of the error.
This manual activity is
tedious and prone to error~\cite{santolucito2018statically}.
As a
single build log can be over 50 megabytes large~\cite{beller2017oops},
finding the small chunk of information in it which is linked to the error, is akin to searching for a
needle in a haystack.
In addition to helping understand and fix build failures, the information richness of build logs enables a plethora of other applications: with the help of
build logs, we can better understand and categorize CI
failures and compile
errors~\cite{islam2017insights,seo2014programmers}, differentiate
testing practices~\cite{orellana2017differences,vassallo2017a-tale},
train classifiers to predict the build outcome and
duration~\cite{ni2017cost,bisong2017built,machalica2019predictive}, or
investigate static analysis tools in CI
builds~\cite{zampetti2017open}.

However, manual analysis of build logs does not
scale to the number of build logs and amount of information that such
sophisticated applications require. Moreover, being able to
automatically extract relevant pieces of information from build logs can also help
developers debug a broken build more quickly without having to browse the entire build log.

@@

Despite its central role for helping developers and the enablement of
on-ward applications of build logs, the process of how to
automatically analyze build logs has thus far received little
attention as a separate research topic, bringing us to our first
research question:
\begin{simplebox}[minipage boxed title*=-5cm]{\textbf{Research Question
1}}
What are the state-of-the-research techniques to analyze build logs?
\end{simplebox}

To answer this question, we start our investigation by performing an extensive literature study (\Cref{fig:overview}).
It shows that
less than half of the 61 works which rely on build log analysis
describe in detail how they performed it, and even fewer
made their implementation available.
The literature
survey identifies three prevalent strategies to
analyze build logs: Using 1) a set of
manually curated regular expressions,
similar to TravisTorrent~\cite{beller2017oops},
2) information
retrieval techniques,
% TODO: Can you shortly explain the technique here?
and 3) a list of search keywords
to identify regions of interest in the build log.

Even though these three approaches to retrieve information chunks from
build logs have different strengths and weaknesses, the literature
survey shows that their use often seems arbitrary.
As a consequence,
in RQ2 we ask:

@@

To answer this question, in the second step of our study
(\Cref{fig:overview}), we build and compare three
implementations that each stand protoypical for one of the idenitied
strategies to analyze build logs:

\begin{itemize}
  \item \textbf{Program Synthesis by Example (referred to as PBE)}
  Based on manually supplied example pairs of build logs and chunks, PBE automatically
  synthesizes
  a regular expression extracting the given chunks within the build logs.
  \item \textbf{Common Text Similarity (CTS)}
  % TODO what is the vector space model
  Using the Vector Space Model, CTS selects lines of a log which are
  most similar to the lines present in manually supplied example chunks.
  \item \textbf{Keyword Search (KWS)}
  % TODO what is an ad-hoc method?
  Ad-hoc method for finding fitting passages by searching the whole
  text of a build log for
  the occurrence of specific trigger words.
\end{itemize}

In the third step, we conduct an empirical study to gauge the performance of
these chunk retrieval techniques in
real-world projects under a variety of conditions.
Our study is based on the \emph{LogChunks} data
set~\cite{brandt2020logchunks}, which encompasses build logs from
80 popular open-source projects and a broad range of different
build tools and programming languages.

Finally, we develop a set of guidelines (step 4) on how to choose a technique given a goal.
We recommend PBE for use cases where the desired information is always
represented in the same structural way and high confidence in
precision and recall of the chunk retrieval is required.
Results
produced by PBE are suited for automatic on-ward processing.
CTS is
well-suited when the representation of the desired information varies
slightly and the output of the chunk retrieval is further processed by
a human.
In cases where the textual representation of the desired
information in the log is unpredictable or varies greatly, KWS seems
to be the best choice.
However, its low precision---it extracts a
context of multiple lines around a finding---makes it generally
unsuited for automatic on-ward processing and instead requires a human
to further inspect and interpret the output chunk. A hybrid-solution combining the three techniques might be the best allround solution.

In short, this article contributes
\begin{itemize}
\item the first systematic study of the emerging field of build log analysis.
\item three prototypical open-sourced implementations of the most promising chunk retrieval techniques.
\item a large-scale empirical study comparing the three techniques.
\item systematic guidelines on when to use which technique.
\item a detailed replication package~\cite{brandt2020chunk-replication}.
\end{itemize}

@@

In this section, we survey the emerging field of build log
analysis. We first start with a differentiation to the related field of general system log analysis. Then, we describe the methodology and results of a systematic mapping study to
determine the state-of-research on build log analysis following the
guidelines by Petersen et
al.~\cite{petersen2008systematic,petersen2015guidelines}. 
We close by summarizing how build logs are commonly
analyzed in literature and discuss the implications of these findings.

@@

\subsection{Distinction from System Log Analysis}
\label{sec:system-log-analysis}
A field related to build log analysis is the processing of system log
files produced during runtime of the system. The first difference here is that build logs are produced during runtime of the build of the system, not the system itself.
In terms of the contents of the logs, the main difference is that system logs are fundamentally structured
through events.
Each line in a log file represents one event with a
set of fields: timestamp, verbosity level and raw message
content~\cite{he2017towards}.
\Cref{lst:system-log} shows
example lines from a system log.

The first goal in analyzing system log files is generally to
separate constant and variable parts within a log
message~\cite{nagappan2010abstracting,he2017towards}.
Next, the log
messages are clustered into log events, unifying messages with
identical constant parts and varying parameters. Then, a log parser works on this intermediate representation.
Its output is an ordered list of timed events and their corresponding
parameter values~\cite{he2016evaluation}.
This structured log can serve as input to various machine learning and data mining processes.

In principle, one can apply the techniques developed for system log analysis
to parts of the build logs that have this same structure.
% TODO this explanation is not understandable. It needs to better embedded in the surrounding text
One example is comparing execution traces to reference
traces of intended behavior to detect anomalies.
Amar et
al.~\cite{amar2019mining} employed a similar approach to detect
relevant lines in build logs.
However, almost all of the techniques developed to analyze system logs
leverage their inherent event structure.
Build logs generally lack this structure and thus require different bespoke approaches.
% In this article, we focus on extracting a single specified information
% from the build log as a whole with chunk retrieval techniques.
% Chunk
% retrieval techniques are used as a part of log parsing to retrieve the
% values of variable parts in a log message, e.g.\ by using regular
% expressions~\cite{nagappan2010abstracting,xu2009detecting}.

% TODO In general, we should reference petersen2015guidelines a bit
% more. I.e., following the suggestions by Petersen or something, to
% show how we followed their protocol. This comment applies to the rest of this section

\subsection{Study Selection}
\Cref{fig:lit-survey} presents an overview of our study selection
process.
Since the field of build log analysis is relatively young, we
surveyed as broad a population of scientific material as possible.
We did not limit our sources to specific Software Engineering venues~\cite{petersen2015guidelines}, but survey all sources
monitored by Google Scholar, including Bachelor's and Master's
theses and academic slide decks.
We use Google Scholar as opposed to Scopus or
other databases because it is the search engine for academic articles with
the broadest and most current index, including preprints.

An initial Google Scholar search for variations of ``build log''
returned more than 2.2 million results (2020-03-15), many of which stem from unrelated fields such as construction and wood working.
We refined our search criteria to exclude such obvious
non-related fields and also to exclude works on systems
or event logs (see \Cref{sec:system-log-analysis}).
This left us with 1,552 search results
(2020-03-15).
Our replication package documents the full search queries~\cite{brandt2020chunk-replication}.
To make handling so many search results feasible, we employed a
three-pass filtering strategy:

First, we filtered articles based on (a) their title and (b) the
contextual information Google Scholar displayed.
We included works at large that seemed to bear a resemblence to software engineering and building software. We excluded wroks written in a language unintelligible to the authors (i.e., not English or German), which disregarded fewer than 1\% of search
hits.
We then unified
the results of the four search queries based on the link as the
identifying element, removing 60 articles that appeared in more than
one search query.
We removed further duplicates based on the title and were
left with 256 (16\%) of the original search results.
% removed duplicates of data extraction pass already here
With these 256 remaining articles, we followed a backward snowball
sampling for related work.
If an article referenced a new work in the context
of build log analysis, we added the new one to our literature set.
This added five articles (eight before
duplicate elimination), leading to a total of 261
works.

Subsequently, we performed a second, finer filter phase by downloading the
full text of the works, reading their abstracts, and searching the full text
for the occurrence of ``log.''
If a work showed traces of
working with build logs, we included it in our survey.
In total, we
were left with 87 works after the second filtering phases.

\subsection{Article Classification}
Having trimmed down the set of articles by 95\%, we investigated the remaining ones in depth.
We wanted to characterize how research uses build logs.

% TODO (resolved): RQs mentioned before explicitly stated
Specifically, we were interested in
\begin{itemize}
  \item what information is retrieved from the logs and
  what the retrieved information is used for (\textbf{RQ1.1})
  \item which techniques are used to process the build logs,
  in how much detail they are described, and whether they are published as tools (\textbf{RQ1.2})
  \item what kind of build logs are used and the origin of these
  logs.
(\textbf{RQ1.3})
\end{itemize}

This lead us to the following research questions:
\begin{simplebox}[attach boxed title to top center={yshift=-6mm}]
{\textbf{RQ1:} What are the state-of-the-research
techniques to analyze build logs?}
\begin{itemize}[leftmargin=2cm]
  \item[\textbf{RQ1.1:}] Which information is targeted?
  \item[\textbf{RQ1.2:}] Which techniques are used?
  \item[\textbf{RQ1.3:}] Which build logs are analyzed?
\end{itemize}
\end{simplebox}

Since we had been careful to include rather than exclude borderline works, this deeper evaluation excluded another 15 articles from the 87 works left after the second filter phase.
Multiple articles were extensions of others.
Of these, we only considered the earliest article which reported on
the processing of build logs.
This excluded another 11 articles.
In sum, we extracted data from 61 articles.

To guide the data extraction we created a template. The template form contained questions to answer RQ 1.1 through RQ 1.3, for example ``Is the technique explained in sufficient detail?'' and ``What is the source of the build logs?'' As recommended by Petersen et
al.~\cite{petersen2015guidelines}, we started the mapping study with a pilot study:
both authors extracted  data from the same five articles and held a
consensus meeting to unify their understanding
of the data extraction template.
We divided the remaining 46 articles evenly between the two authors and inspected them as closely as necessary to fill out the template with confidence,
starting from occurrences of ``log'' or ``build'' within the article's text. For questions that did not require a boolean answer, we allowed assigning multiple tags or categories as the first part (``preparation phase'') of a virtual open card sorting~\cite{zimmermann2016card}. 
% TODO: add more concrete example, such as use of retrieved x: developer feedback, build data set
Once we had classified all works, we went into the second phase (``execution phase'') in which we found synonyms, and grouped tags together.

%The full data extraction template and our results of this step can be
%found in our replication package~\cite{brandt2020chunk-replication}.

% TODO (resolved): RQs just appeared unconnected out of the blue

@@

\subsection{Results}
In this section, we present the results of the analysis phase of the
literature mapping along RQ 1.1 through RQ 1.3.

% TODO repeat the question in the subsection title?
\subsubsection{Retrieved Information (RQ 1.1)}
% TODO General ref to Figure 4 missing: We answer this question along data in Figure 4, and ... ?

The precise information targeted in the build logs varied from article to article.
\Cref{fig:litsur:info_target} gives an overview over the frequency of the targeted information. Most prominent was the search for errors (39\%), followed by extracting the executed or the failing
tests (both 11\%).
There were also numerous works which targeted more specialized
information, such as the environments used during the
build~\cite{zolfagharinia2017not}, packages
installed~\cite{selberg2012use}, or hints to source files which 
relate to build failures~\cite{ren2018automated}.

% TODO It is unclear where this comes from. Also, wouldn't it perhaps be better to continue with Figure 4.b directly?
In most cases (51\%), a chunk was retrieved from the log.
Build logs were scanned for compilation
errors~\cite{clemencic2014new} or
the duration of tasks within the build~\cite{zhang2016android}.
36\% of the articles classified the build logs into categories.
For instance, distinguishing failures caused by developers from
failures caused
by infrastructure errors~\cite{lindqvist2019detection} or
whether projects use automated static analysis
tools~\cite{kavaler2019tool}.
Other types of information (e.g.,\ counting
lines, summarization) were only retrieved
in one case each.

The majority of articles (46\%) used the outcome of their build log
analysis for research purposes, as shown in \Cref{fig:litsur:use}.
% TOOD What does this mean? Give an example: to drive further research, for example to categorize the types of compile errors encountered at Google~\cite{}.
Further uses were giving feedback to developers (10\%) or automatic
on-ward processing of the retrieved information (8\%)
% TODO for example to ...
.

% TODO repeat the question in the subsection title?
\subsubsection{Retrieval Technique (RQ 1.2)}
\Cref{tab:litsur:techniques} presents the techniques that the articles
% TODO: lets use analyzing build logs canonically? Please search text to eliminate occurrences of processing?
described for processing build logs.
% TODO This is weird. Last number we have from above is 51.
43 of the 61 articles we inspected mentioned how they are analyzing build
logs.
However, only 16 of these described their method in detail.

The most mentioned method (34\%) was using a parser, where we also
included
rather imprecise statements like ``we parse the build logs.''
% TODO include citation?
% TOOD maybe make the point here that it is questionable whether this includes a comprehensive tokenizer, lexer etc. based solution and that parsing in many cases really meant regexp. If we could figure that out, then we would assign it to regexp instead.
25\% of the articles described the use of regular expressions and 20\%
inspected the logs manually.
Another 8\% employed machine learning such as natural language
processing or information retrieval techniques.
For example,
Seo et al.~\cite{seo2014programmers} developed a custom
parser to classify error messages, while Vassallo et
al.~\cite{vassallo2017a-tale} analyzed build logs with regular
expressions.
Ghaleb et al.~\cite{ghaleb2019studying} used a compound approach.
They started with manual categorizing build logs and selecting
search strings that identify the target category.
Based on the search strings they created a script that automatically
classifies the remaining logs.

Only 25\% of the articles claimed their implementation is available. Moreover, we
saw some reuse of the few available build log analysis tools.
Five articles employed the tools used to create
\emph{TravisTorrent}~\cite{beller2017travistorrent,beller2017oops,
orellana2017differences,zhao2018comparing} or
an extended version of them~\cite{rott2019empirische,
shi2018evaluating} (based on regular expressions),
two used the
\emph{Maven Log Analyzer}~\cite{macho2018automatically,gallaba2018noise}
% TODO: based on which technique?,
another two the
\emph{MAKAO}~\cite{wen2018blimp,adams2007design,adams2007makao} tool
% TODO: based on which technique?
.

\subsubsection{Build Logs}
% TODO using passive where it is unneccessary
The distribution of supported build tools is shown in
\Cref{fig:litsur:log_producer}.
Of the articles we surveyed, 30\% analyzed build logs produced by
Maven~\cite{maven2019website}, 16\% analyzed logs produced by
Gradle~\cite{gradle2020website}
and 10\% analyzed logs produced by Ant~\cite{ant2020website}.

% TODO remove first part of this paragraph?
Sometimes this limitation to a specific build tool is motivated by
aspects separate from the build log analysis.
For example, Shi et al.~\cite{shi2018evaluating} focussed on Maven
build logs because they also chose the PIT tool to calculate coverage
and mutation score.
A larger amount of the articles within our survey, however, limit
themselves to few build tools because their log format varies and
therefore ``parsers must be specialized to each build and test
framework''~\cite{tomassi2019bugswarm}.

Only 7\% of the proposed methods claimed to be language-agnostic, while
several described they are covering multiple source build tools.
The majority of articles (42\%) targeted logs from a source that no second article
targeted.

% TODO death by passive
The main platform that build logs were collected from is Travis
CI~\cite{travisci2019webpage},
which was used in 38\% cases.
18\% of the articles processed logs from TravisTorrent and 8\% included
build logs from industrial projects.

% perhaps remove?
The number of logs analyzed varied greatly and we were not always able to
extract it confidently.
Articles reported processing from two up to 122 million logs.

31\% of the surveyed articles claimed to have published their data to
enable further research.
However, we found no clear reuse of build log data sets with the
exception of TravisTorrent~\cite{beller2017travistorrent}, which 
Vassallo et al.~\cite{vassallo2017a-tale},
Ghaleb et al.~\cite{ghaleb2019studying}, and several other studies reused.

% TODO A and B should possibly be overlapping, otherwise there is an area where neither A nor B makes sense?
\begin{figure}[tbhp]
		\centering
		\includegraphics[width=\columnwidth, trim={1.1cm 0.4cm
		1.5cm 0.5cm},
		clip]{img/lit-sur/log_producer_annotated.pdf}
		\caption{Frequency of supported log producers.}
		\label{fig:litsur:log_producer}
\end{figure}

\subsection{Discussion}
\label{sec:lit-sur:discussion}

% TODO: this section is crucial. maybe we can highlight important findings, by categorizing them either as further subsubsections or making it bold in the text?
% TODO: find values for X and Y
Our literature survey shows that despite its relatively young age
(first occurrence of a paper on the topic in X), \textbf{build log analysis is
a widely established technique} in the literature, with Y articles on it.

% TODO: this reads very hands-off and researchy.
% I don't understand the first sentence and would simply cut it.
We saw that various researchers process build logs, often because the logs  provide more or the necessary
information to the researchers~\cite{ren2018automated}.  \textbf{Most log analysis is automated.} Automated approaches are necessary as many studies target a large
number of builds. Articles
which explicitly mention a manual analysis of the logs, mostly
restrict their work to few logs or take a ``representative sample to
make manual analysis feasible''~\cite{zolfagharinia2017not}.
% The chunk retrieval techniques we investigate within this article are
% automated, apart from their configuration through training examples.

% TODO: draw a conclusion together with figure 5: if reuse were higher, lots of effort could be saved?
% discussion about popularity of replication packages adds nothing to this paper.
We observed very little reuse of build log analysis tools.
This can stem from the fact that publishing replication packages and
research prototypes became popular only in recent years or from the
high specialization of the developed tools in regards to supported
build environments and targeted information.
